---
title: "ESG vs Conventional Indices: Time Series Analysis"
author: "Alison Barnard and Anya Low"
date: "2026"
output: pdf_document
---

## 1. Introduction

The concept of ESG investing — incorporating Environmental, Social, and 
Governance criteria into investment decisions — has grown substantially 
over the past decade. Between 2020 and 2021, ESG-labelled funds attracted 
record inflows as institutional and retail investors increasingly 
prioritised sustainability alongside financial returns. However, the period 
that followed presented a markedly different environment. Rising interest 
rates from 2022 onwards disproportionately affected growth-oriented and 
ESG-heavy sectors such as technology and renewable energy, while a growing 
wave of institutional scepticism toward ESG products contributed to 
significant outflows from ESG funds.

This shifting landscape raises an important empirical question: did ESG 
indices actually behave differently from their conventional counterparts 
during this period, and if so, can any differences be explained by the 
interest rate environment?

This paper investigates the return and volatility dynamics of ESG-screened 
equity indices relative to conventional benchmarks across two market 
pairs — one US and one global — over the period January 2020 to December 
2024. We apply a structured time series framework including descriptive 
analysis, stationarity testing, autocorrelation analysis, GARCH volatility 
modelling, and regime analysis across pre- and post-2022 subsamples.

---

## 2. Literature Review

Research on the financial performance of ESG investments relative to 
conventional benchmarks has grown substantially alongside the rise of 
sustainable investing. Friede, Busch and Bassen (2015) conducted a 
meta-analysis of over 2,000 empirical studies and found that the large 
majority reported a non-negative relationship between ESG criteria and 
corporate financial performance, suggesting that ESG investing does not 
systematically destroy value.

More recent work has focused on the dynamic relationship between ESG 
and conventional indices during periods of market stress. Ashraf (2020)
demonstrated that stock markets responded negatively to 
rising COVID-19 case numbers across 64 countries, with the strongest 
negative reaction occurring in the early stages of the outbreak.

Albuquerque et al.(2020) further showed that stocks with higher environmental
and social ratings exhibited significantly higher returns and lower volatility 
during the COVID-19 market crash of Q1 2020, suggesting that ESG 
screening may confer resilience during periods of acute market stress.

Albuquerque et al. (2020) attribute this resilience to stronger customer 
and investor loyalty among ESG firms, which dampens both revenue 
volatility and selling pressure during market stress.

However, the relationship between ESG performance and the interest rate 
environment has received comparatively less attention. ESG indices tend 
to be structurally overweight in growth and technology stocks and 
underweight in energy and value stocks — sectors that typically 
outperform during rising rate environments (Pastor, Stambaugh and 
Taylor, 2022). This structural tilt suggests that rising interest rates 
may systematically disadvantage ESG indices relative to conventional 
benchmarks, a hypothesis we test empirically in this paper.

---

## 3. Research Question and Hypotheses

Our central research question is:

> **Did the return and volatility dynamics of ESG indices change relative 
> to conventional indices during the 2022–2024 rising interest rate period?**

We explore this through the following hypotheses:

**H1:** ESG indices exhibit different return behaviour compared to 
conventional indices during the 2022–2024 period.

**H2:** ESG indices exhibit different volatility dynamics compared to 
conventional indices.

**H3:** The relationship between ESG and conventional indices differs 
between the pre-2022 and post-2022 subsamples.

---

# 3. Data Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r install-packages, eval=FALSE}
install.packages("quantmod")
install.packages("zoo")
install.packages("xts")
install.packages("ggplot2")
install.packages("moments")
install.packages("reshape2")
install.packages("kableExtra")
install.packages("dplyr")
install.packages("rugarch")
```

```{r libraries}
library(quantmod)
library(zoo)
library(xts)
library(ggplot2)
library(moments)
library(reshape2)
library(kableExtra)
```

```{r parameters}
start_date <- "2020-01-01"
end_date   <- "2024-12-31"
```

We created the date parameters for data retrieval from the Yahoo website, which can interact with R through "quantmod" package.

```{r pull-equity-data}
getSymbols("ESGU",    from = start_date, to = end_date, src = "yahoo")
getSymbols("^GSPC",   from = start_date, to = end_date, src = "yahoo")
getSymbols("IWDA.AS", from = start_date, to = end_date, src = "yahoo")
getSymbols("SUWS.L",  from = start_date, to = end_date, src = "yahoo")
```

```{r pull-rate-data}
getSymbols("DGS10", src = "FRED", from = start_date, to = end_date)
```

```{r extract-prices}
esgu <- Ad(ESGU)
gspc <- Ad(GSPC)
iwda <- Ad(IWDA.AS)
suws <- Ad(SUWS.L)
rate <- DGS10
```

```{r check-data, include=FALSE}
head(esgu)
head(gspc)
head(iwda)
head(suws)
head(rate)
```
## 4. Data and Methodology

### 4.1 Data Sources and Collection

The dataset comprises five daily time series covering the period from 
2 January 2020 to 31 December 2024, yielding 1,304 trading day 
observations after cleaning. Two pairs of ESG and conventional indices 
are used: the iShares ESG Aware MSCI USA ETF (ESGU) versus the S&P 500 
(^GSPC) for the US market, and the iShares MSCI World ESG Leaders ETF 
(SUWS.L) versus the iShares Core MSCI World ETF (IWDA.AS) for the global 
market. The 10-year US Treasury yield (DGS10) is included as a proxy for 
the interest rate environment. All equity data is sourced via the Yahoo 
Finance API using the quantmod package in R. Interest rate data is 
sourced from the Federal Reserve Economic Data (FRED) platform.

The series are denominated in different currencies and price scales. 
This is expected and does not affect the analysis, as all subsequent 
calculations are based on log returns rather than price levels.

### 4.2 Data Cleaning and Preparation

```{r merge-and-rename}
all_data <- merge(esgu, gspc, iwda, suws, rate)
colnames(all_data) <- c("ESGU", "GSPC", "IWDA", "SUWS", "Rate_10Y")
```

After merging, missing values were identified across all series due to 
differences in trading calendars between US and European markets. US 
markets observe different public holidays to Dutch and UK markets, 
resulting in non-overlapping non-trading days. Missing values were 
handled by carrying the last available observation forward using 
na.locf(), which is standard practice in financial time series analysis.

```{r check-nas, include=FALSE}
colSums(is.na(all_data))
```

```{r clean-data}
all_data <- na.locf(all_data)
all_data <- na.omit(all_data)
```

```{r check-quality, include=FALSE}
colSums(all_data <= 0, na.rm = TRUE)
index(all_data)[1]
index(all_data)[nrow(all_data)]
nrow(all_data)
```

Following cleaning, the dataset runs from 2 January 2020 to 31 December 
2024, comprising 1,304 trading days — consistent with approximately 260 
trading days per year over five years. No zero or negative prices were 
identified, confirming data integrity.

Log returns are computed for all equity series using the formula:

$$r_t = \ln(Y_t) - \ln(Y_{t-1})$$

Log returns are preferred over simple percentage returns as they are 
additive over time and more closely approximate a normal distribution, 
making them better suited for time series modelling. For the interest 
rate series, the first difference of the yield level is used rather 
than log returns, as the rate is already expressed as a percentage.

```{r compute-returns}
log_returns <- diff(log(all_data))
log_returns <- na.omit(log_returns)
rate_diff   <- diff(all_data$Rate_10Y)
rate_diff   <- na.omit(rate_diff)
```

```{r save-data}
dir.create("data", showWarnings = FALSE)
write.csv(as.data.frame(all_data),    "data/prices.csv",      row.names = TRUE)
write.csv(as.data.frame(log_returns), "data/log_returns.csv", row.names = TRUE)
write.csv(as.data.frame(rate_diff),   "data/rate_diff.csv",   row.names = TRUE)
```

### 4.3 Analytical Methods

The analysis follows a structured time series pipeline. First, 
descriptive analysis is conducted to compare return distributions, 
rolling volatility, and drawdowns across ESG and conventional series. 
Second, Augmented Dickey-Fuller (ADF) tests are applied to confirm 
stationarity of the transformed series before modelling. Third, 
autocorrelation (ACF) and partial autocorrelation (PACF) functions are 
examined to assess dependence structures. Fourth, GARCH(1,1) models are 
estimated to capture volatility clustering and persistence within each 
series. Fifth, a regime analysis compares pre-2022 and post-2022 
subsamples to evaluate whether return and volatility dynamics changed 
with the interest rate environment. Finally, a VAR model is used to 
examine dynamic interactions between the ESG-conventional performance 
spread and the interest rate series.

## 5. Results

### 5.1 Descriptive Analysis

#### 5.1.1 Summary Statistics

The table below presents key distributional statistics for the log 
return series of all four equity indices over the full sample period 
from January 2020 to December 2024.

```{r descriptive-stats}
library(moments)

desc_stats <- data.frame(
  Series   = c("ESGU", "GSPC", "IWDA", "SUWS"),
  Mean     = c(0.00050, 0.00046, 0.00046, 0.00038),
  Std_Dev  = c(0.01343, 0.01326, 0.01052, 0.01159),
  Skewness = c(-0.806, -0.832, -0.926, -0.656),
  Kurtosis = c(17.436, 17.780, 13.639, 10.573),
  Min      = c(-0.1279, -0.1277, -0.0935, -0.1007),
  Max      = c(0.0925, 0.0897, 0.0805, 0.0773)
)

knitr::kable(desc_stats,
             col.names = c("Series", "Mean", "Std Dev", 
                           "Skewness", "Kurtosis", "Min", "Max"),
             caption = "Table 1: Descriptive Statistics of Daily Log Returns (Jan 2020 – Dec 2024)",
             digits = 5,
             align = c("l", "r", "r", "r", "r", "r", "r"),
             booktabs = TRUE) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                full_width = FALSE) |>
  row_spec(0, bold = TRUE)
```

All four series exhibit negative skewness, indicating that large 
negative returns occur more frequently than large positive returns — 
a common finding in equity return data. The kurtosis values exceed 3 
for all series, indicating fat tails relative to the normal distribution. 
This suggests that extreme return events occur more frequently than a 
normal distribution would predict, which is consistent with the 
turbulent market conditions observed during the sample period including 
the COVID-19 shock in 2020 and the rate hiking cycle beginning in 2022.

#### 5.1.2 Price Level Visualisation

```{r price-plot}
library(ggplot2)
library(reshape2)

# Convert to data frame for ggplot
prices_df <- data.frame(
  Date = index(all_data),
  ESGU = as.numeric(all_data$ESGU),
  GSPC = as.numeric(all_data$GSPC),
  IWDA = as.numeric(all_data$IWDA),
  SUWS = as.numeric(all_data$SUWS)
)

# Normalise to 100 at start so all series are comparable
prices_df[, 2:5] <- sweep(prices_df[, 2:5], 2, 
                           as.numeric(prices_df[1, 2:5]), "/") * 100

prices_long <- melt(prices_df, id.vars = "Date", 
                    variable.name = "Series", value.name = "Price")

ggplot(prices_long, aes(x = Date, y = Price, colour = Series)) +
  geom_line() +
  geom_vline(xintercept = as.Date("2022-01-01"), 
             linetype = "dashed", colour = "black") +
  annotate("text", x = as.Date("2022-03-01"), y = 80, 
           label = "Rate hiking begins", size = 3, hjust = 0) +
  labs(title = "Normalised Price Series (Base = 100, January 2020)",
       x = "Date", y = "Indexed Price") +
  theme_minimal()
```

#### 5.1.3 Log Returns Visualisation

```{r returns-plot}
returns_df <- data.frame(
  Date = index(log_returns),
  ESGU = as.numeric(log_returns$ESGU),
  GSPC = as.numeric(log_returns$GSPC),
  IWDA = as.numeric(log_returns$IWDA),
  SUWS = as.numeric(log_returns$SUWS)
)

returns_long <- melt(returns_df, id.vars = "Date",
                     variable.name = "Series", value.name = "Return")

ggplot(returns_long, aes(x = Date, y = Return, colour = Series)) +
  geom_line(alpha = 0.7) +
  geom_vline(xintercept = as.Date("2022-01-01"), 
             linetype = "dashed", colour = "black") +
  labs(title = "Daily Log Returns",
       x = "Date", y = "Log Return") +
  theme_minimal()
```
As shown in Figure \@ref(fig:returns-plot), the daily log returns for all four indices over the full sample period. The most prominent feature is the sharp clustering of large 
positive and negative returns in early 2020, coinciding with the onset of 
the COVID-19 pandemic. Returns during this period reach magnitudes of 
approximately ±10%, far exceeding typical daily variation. This volatility 
cluster dissipates over the course of 2020 and returns remain relatively 
subdued through 2021.

Following the onset of the rate-hiking cycle in early 2022 (dashed vertical 
line), a modest but visible increase in return variability is observed across 
all series, consistent with heightened macroeconomic uncertainty during this 
period. Crucially, the four series track each other closely throughout, with 
no obvious divergence between ESG and conventional indices — suggesting that 
ESG screening does not substantially alter the return dynamics of the 
underlying market exposure.

#### 5.1.4 Rolling Volatility

Rolling volatility measures how the riskiness of each series changes 
over time. We compute a 30-day rolling standard deviation of log returns 
for each series to visualise volatility clustering across the sample period.

```{r rolling-volatility}
library(zoo)

roll_vol <- data.frame(
  Date = index(log_returns),
  ESGU = rollapply(log_returns$ESGU, width = 30, FUN = sd, fill = NA),
  GSPC = rollapply(log_returns$GSPC, width = 30, FUN = sd, fill = NA),
  IWDA = rollapply(log_returns$IWDA, width = 30, FUN = sd, fill = NA),
  SUWS = rollapply(log_returns$SUWS, width = 30, FUN = sd, fill = NA)
)

roll_vol_long <- melt(roll_vol, id.vars = "Date",
                      variable.name = "Series", value.name = "Volatility")

ggplot(roll_vol_long, aes(x = Date, y = Volatility, colour = Series)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.Date("2022-01-01"), 
             linetype = "dashed", colour = "black") +
  annotate("text", x = as.Date("2022-03-01"), y = 0.025, 
           label = "Rate hiking begins", size = 3, hjust = 0) +
  labs(title = "30-Day Rolling Volatility of Log Returns",
       x = "Date", y = "Rolling Std Dev") +
  theme_minimal()
```

The rolling volatility chart reveals two clear spikes — the first 
corresponding to the COVID-19 market shock in March 2020, and the 
second to the onset of the rate hiking cycle in early 2022. Both 
ESG and conventional series show similar volatility patterns, though 
the relative magnitude of these spikes across pairs will be examined 
more formally in the regime analysis in Section 5.5.

### 5.2 Stationarity Testing

Before applying time series models, the Augmented Dickey-Fuller (ADF) 
test is applied to confirm the stationarity properties of each series. 
The null hypothesis of the ADF test is that the series is non-stationary. 
A p-value below 0.05 allows us to reject the null and conclude the series 
is stationary.

#### 5.2.1 ADF Test on Price Levels


```{r stationarity-levels, echo=FALSE, message=FALSE, warning=FALSE}
library(tseries)
library(knitr)
library(kableExtra)

run_adf <- function(series, name) {
  result <- adf.test(na.omit(series))
  data.frame(
    Series     = name,
    `ADF Stat` = round(result$statistic, 4),
    `Lag Order` = result$parameter,
    `p-value`  = round(result$p.value, 4),
    Result     = ifelse(result$p.value < 0.05, "Stationary", "Non-stationary"),
    check.names = FALSE
  )
}

levels_table <- rbind(
  run_adf(all_data$ESGU,     "ESGU"),
  run_adf(all_data$GSPC,     "GSPC"),
  run_adf(all_data$IWDA,     "IWDA"),
  run_adf(all_data$SUWS,     "SUWS"),
  run_adf(all_data$Rate_10Y, "Rate_10Y")
)

kable(levels_table, booktabs = TRUE, align = "lrrrr",
      caption = "ADF Test Results — Price Levels") |>
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 10) |>
  column_spec(1, bold = TRUE) |>
  column_spec(5, bold = TRUE) |>
  footnote(general = "H0: series has a unit root (non-stationary). Rejection at p < 0.05 indicates stationarity.",
           general_title = "Note: ", footnote_as_chunk = TRUE)
```

As expected, the ADF test results for all price level series return 
p-values above 0.05, confirming that the raw price series are 
non-stationary. This means we cannot use price levels directly in our 
models and must transform the series before proceeding.

#### 5.2.2 ADF Test on Log Returns

```{r stationarity-returns, echo=FALSE, message=FALSE, warning=FALSE}
library(tseries)
library(knitr)
library(kableExtra)

returns_table <- rbind(
  run_adf(log_returns$ESGU,  "ESGU"),
  run_adf(log_returns$GSPC,  "GSPC"),
  run_adf(log_returns$IWDA,  "IWDA"),
  run_adf(log_returns$SUWS,  "SUWS"),
  run_adf(rate_diff,         "Rate_10Y (differenced)")
)

kable(returns_table, booktabs = TRUE, align = "lrrrr",
      caption = "ADF Test Results — Log Returns") |>
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 10) |>
  column_spec(1, bold = TRUE) |>
  column_spec(5, bold = TRUE) |>
  footnote(general = "H0: series has a unit root (non-stationary). All return series reject H0 at p < 0.05, confirming stationarity.",
           general_title = "Note: ", footnote_as_chunk = TRUE)
```

The ADF test results for all log return series return p-values below 
0.05, confirming that the transformed series are stationary and suitable 
for use in subsequent modelling steps. For the interest rate series, the 
first difference of the yield level is used rather than log returns, as 
the rate is already expressed as a percentage.

#### 5.2.3 Summary of Stationarity Results

```{r stationarity-summary}
stationarity_summary <- data.frame(
  Series = c("ESGU (price)", "GSPC (price)", "IWDA (price)", 
             "SUWS (price)", "Rate 10Y (level)",
             "ESGU (returns)", "GSPC (returns)", 
             "IWDA (returns)", "SUWS (returns)"),
  DF_Stat = c(-1.606, -1.756, -2.283, -2.194, -3.139,
              -10.347, -10.422, -9.424, -10.110),
  P_Value = c(0.745, 0.681, 0.459, 0.496, 0.098,
              0.010, 0.010, 0.010, 0.010),
  Stationary = c("No", "No", "No", "No", "No",
                 "Yes", "Yes", "Yes", "Yes")
)

knitr::kable(stationarity_summary,
             col.names = c("Series", "Dickey-Fuller Statistic", 
                           "p-value", "Stationary"),
             caption = "Table 2: ADF Test Results — Price Levels and Log Returns",
             digits = 3,
             align = c("l", "r", "r", "c"),
             booktabs = TRUE) |>
  kable_styling(latex_options = c("hold_position"),
                full_width = FALSE) |>
  row_spec(0, bold = TRUE) |>
  row_spec(5, extra_latex_after = "\\midrule")
```

### 5.4 Volatility Modelling (GARCH)

While rolling volatility provides a useful visual summary of how 
volatility changes over time, it does not formally model the statistical 
properties of volatility. To do this we fit a GARCH(1,1) model to the 
log return series of each index. The GARCH(1,1) model captures two 
well-documented features of financial return data: volatility clustering 
(periods of high volatility tend to be followed by further high 
volatility) and mean reversion (volatility eventually returns to its 
long-run average).

The GARCH(1,1) model is specified as follows:

$$\sigma^2_t = \omega + \alpha \epsilon^2_{t-1} + \beta \sigma^2_{t-1}$$

Where $\sigma^2_t$ is the conditional variance at time $t$, 
$\epsilon^2_{t-1}$ is the squared return shock from the previous period 
(the ARCH term), and $\sigma^2_{t-1}$ is the conditional variance from 
the previous period (the GARCH term). The parameter $\alpha$ captures 
how strongly volatility reacts to new shocks, and $\beta$ captures how 
persistent volatility is over time. A sum of $\alpha + \beta$ close to 1 
indicates high volatility persistence.

```{r garch-models, include=FALSE, message=FALSE, warning=FALSE}
library(rugarch)
library(xts)

garch_spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model     = list(armaOrder = c(0, 0), include.mean = TRUE),
  distribution.model = "norm"
)

garch_esgu <- ugarchfit(spec = garch_spec, data = na.omit(log_returns$ESGU),
                        solver = "hybrid", solver.control = list(trace = 0))
garch_gspc <- ugarchfit(spec = garch_spec, data = na.omit(log_returns$GSPC),
                        solver = "hybrid", solver.control = list(trace = 0))
garch_iwda <- ugarchfit(spec = garch_spec, data = na.omit(log_returns$IWDA),
                        solver = "hybrid", solver.control = list(trace = 0))
garch_suws <- ugarchfit(spec = garch_spec, data = na.omit(log_returns$SUWS),
                        solver = "hybrid", solver.control = list(trace = 0))
```

```{r garch-summary-table, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(rugarch)

extract_coef <- function(fit, name) {
  cf  <- coef(fit)
  ic  <- infocriteria(fit)
  a   <- cf["alpha1"]
  b   <- cf["beta1"]
  df <- data.frame(
    Series        = name,
    `ω`           = round(cf["omega"],  6),
    `α (ARCH)`    = round(a,            4),
    `β (GARCH)`   = round(b,            4),
    `α + β`       = round(a + b,        4),
    AIC           = round(ic[1],        4),
    BIC           = round(ic[2],        4),
    check.names   = FALSE
  )
  rownames(df) <- NULL    # <-- add this line here
  df                      # this line was already there
}

coef_table <- rbind(
  extract_coef(garch_esgu, "ESGU"),
  extract_coef(garch_gspc, "GSPC (S&P 500)"),
  extract_coef(garch_iwda, "IWDA"),
  extract_coef(garch_suws, "SUWS")
)

kable(coef_table, booktabs = TRUE, align = "lrrrrrr",
      caption = "GARCH(1,1) Parameter Estimates") |>
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 10) |>
  column_spec(1, bold = TRUE) |>
  column_spec(5, bold = TRUE) |>
  footnote(general = "α + β close to 1 indicates high volatility persistence. AIC/BIC are model fit criteria (lower is better).",
           general_title = "Note: ", footnote_as_chunk = TRUE)
```
The GARCH(1,1) estimates reveal highly consistent volatility dynamics 
across all four series. Alpha coefficients range from 0.097 to 0.162, 
indicating that between 10% and 16% of each period's volatility is 
driven by the magnitude of the previous period's return shock. Beta 
coefficients are uniformly high, ranging from 0.804 to 0.871, 
reflecting strong volatility persistence across all series. The sum 
of alpha and beta — a measure of overall volatility persistence — 
falls between 0.963 and 0.969 for all four indices, close to but 
below unity, confirming that volatility is mean-reverting but does 
so slowly.

Notably, the GARCH parameters are remarkably similar between ESG and 
conventional pairs. For the US pair, ESGU (α = 0.157, β = 0.806) and 
GSPC (α = 0.157, β = 0.806) are virtually identical. For the global 
pair, IWDA (α = 0.162, β = 0.804) shows slightly higher shock 
sensitivity than SUWS (α = 0.097, β = 0.871), though both exhibit 
similarly high persistence. This suggests that ESG screening does not 
materially alter the volatility structure of equity returns, and that 
both ESG and conventional indices are subject to the same underlying 
volatility dynamics during the sample period.

```{r garch-vol-plot}
# Extract and plot conditional volatility for all series
vol_df <- data.frame(
  Date = index(log_returns),
  ESGU = as.numeric(sigma(garch_esgu)),
  GSPC = as.numeric(sigma(garch_gspc)),
  IWDA = as.numeric(sigma(garch_iwda)),
  SUWS = as.numeric(sigma(garch_suws))
)

vol_long <- melt(vol_df, id.vars = "Date",
                 variable.name = "Series",
                 value.name = "Volatility")

ggplot(vol_long, aes(x = Date, y = Volatility, colour = Series)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.Date("2022-01-01"),
             linetype = "dashed", colour = "black") +
  annotate("text", x = as.Date("2022-03-01"), y = 0.045,
           label = "Rate hiking begins", size = 3, hjust = 0) +
  labs(title = "GARCH(1,1) Conditional Volatility",
       x = "Date", y = "Conditional Standard Deviation") +
  theme_minimal()
```

The GARCH(1,1) conditional volatility plot refines the picture provided 
by the rolling volatility analysis. Unlike rolling volatility which uses 
a fixed window, GARCH volatility is model-based and reacts dynamically 
to new information. The two clear volatility spikes correspond to the 
COVID-19 market shock in March 2020 and the onset of the rate hiking 
cycle in early 2022.


```{r var-model}
# VAR model goes here
```

```{r regime-analysis}
# Pre/post 2022 comparison goes here
```


## References

Albuquerque, R., Koskinen, Y., Yang, S. and Zhang, C. (2020) 'Resiliency 
of environmental and social stocks: An analysis of the exogenous 
COVID-19 market crash', Review of Corporate Finance Studies, 9(3), 
pp.593–621.

Ashraf, B.N. (2020) 'Stock markets' reaction to COVID-19: Cases or 
fatalities?', Research in International Business and Finance, 54, 
p.101249.

Friede, G., Busch, T. and Bassen, A. (2015) 'ESG and financial 
performance: aggregated evidence from more than 2000 empirical studies', 
Journal of Sustainable Finance & Investment, 5(4), pp.210–233.

Pastor, L., Stambaugh, R.F. and Taylor, L.A. (2022) 'Dissecting green 
returns', Journal of Financial Economics, 146(2), pp.403–424.